A Spectral Analysis of the Domain Decomposed Adjoint Neumann-Ulam Method

### 2012-11-27 doc/mandc2013/notes/spectral_theory.tex
updates to data and plots for spectral analysis

%%---------------------------------------------------------------------------%%
\section{Numerical Experiments}
To test the relationships developed by the spectral analysis, we form
two simple numerical experiments using the diffusion model problem:
one to measure the length of the random walks as a function of the
iteration matrix Eigenvalues, and one to measure the domain leakage
fraction as a function of the iteration matrix Eigenvalues and the
discretization properties. Before doing this, we verify our
computation of the spectral radius of the iteration matrix by
numerically computing the largest Eigenvalue of the diffusion operator
using an iterative Eigenvalue solver. For this verification, a $100
\times 100$ square grid with $h=0.1$ and non-reentrant current
boundary conditions was used and the absorption cross varied from 0 to
100 while the scattering cross section was fixed at
unity. Figure~\ref{fig:measured_spec_rad} gives the measured spectral
radius of the iteration matrix and the computed spectral radius for
the preconditioned diffusion operator using
Eq~(\ref{eq:iteration_radius}) as function of the absorption to
scattering ratio $(\Sigma_a / \Sigma_s)$. We note here that although
there is excellent agreement between the measured and analytic
results, slight differences likely arise due to the non-reentrant
current condition not being included in the spectral analysis. To
verify this (and serve as a better testing environment), reflecting
boundary conditions must be implemented.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_spec_rad.png}
  \end{center}
  \caption{\textbf{Measured and analytic preconditioned diffusion
      operator spectral radius as a function of the absorption cross
      section to scattering cross section ratio.} \textit{The blue
      data was computed numerically by an Eigensolver while the red
      data was generated by Eq~(\ref{eq:iteration_radius}).}}
  \label{fig:measured_spec_rad}
\end{figure}

\subsection{Random Walk Length}
With the spectral data in hand, we can go about setting up an
experiment to measure the length of the random walks generated by the
adjoint Neumann-Ulam solver. To do this, we again use a $100 \times
100$ square grid with $h=0.1$ and non-reentrant current boundary
conditions was used and the absorption cross varied from 0 to 100
while the scattering cross section was fixed at unity. A weight cutoff
of \sn{1}{-4} was used with 10,000 histories generated by a point
source source of strength 1 in the center of the domain such that the
vacuum boundary conditions will have less of an effect on the length
of the random walks. For each of the histories, the number of
transitions made was tallied to provide an effective value of $k$ for
each history. This value was then averaged over all histories to get a
measured value of $k$ for the particular
operator. Figure~\ref{fig:measured_length} gives these measurements as
well as the analytic result computed by Eq~(\ref{eq:analytic_k}) as a
function of the iteration matrix spectral radius,
$\rho(\ve{H})$. Again, we note here that although there is excellent
agreement between the measured and analytic results, slight
differences likely arise due to the non-reentrant current condition
not being included in the spectral analysis.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_length.png}
  \end{center}
  \caption{\textbf{Measured and analytic random walk length as a
      function of the iteration matrix spectral radius.} \textit{The
      blue data was computed numerically by an adjoint Neumann-Ulam
      implementation while the red data was generated by
      Eq~(\ref{eq:analytic_k}).}}
  \label{fig:measured_length}
\end{figure}

@@

To compute the leakage fraction numerically, over the global domain
\sn{1}{5} histories were sampled from a uniform source of strength
unity. At the start of a stage of histories, the number of histories
starting in the domain was computed and as the stage progressed, the
number of histories that exited the domain was tallied with the ratio
of the two numbers providing a numerical measure for the leakage
fraction. Figure~\ref{fig:measured_length} gives the domain leakage
measurements for the domain in the center of the global grid as well
as the analytic result computed by
Eq~(\ref{eq:analytic_domain_leakage}) as a function of the absorption
cross section. Again, we note good qualitative agreement between the
measured and analytic quantities.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_leakage.png}
  \end{center}
  \caption{\textbf{Measured and analytic domain leakage as a function
      of the absorption cross section.} \textit{The blue data was
      computed numerically by a domain-decomposed adjoint Neumann-Ulam
      implementation while the red data was generated by
      Eq~(\ref{eq:analytic_domain_leakage}).}}
  \label{fig:measured_leakage.}
\end{figure}

>>>>

%%---------------------------------------------------------------------------%%
\section{Numerical Experiments}
To test the relationships developed by the spectral analysis, we form
two simple numerical experiments using the diffusion model problem:
one to measure the length of the random walks as a function of the
iteration matrix Eigenvalues, and one to measure the domain leakage
fraction as a function of the iteration matrix Eigenvalues and the
discretization properties. Before doing this, we verify our
computation of the spectral radius of the iteration matrix by
numerically computing the largest Eigenvalue of the diffusion operator
using an iterative Eigenvalue solver. For this verification, a $100
\times 100$ square grid with $h=0.1$ and non-reentrant current
boundary conditions was used and the absorption cross varied from 0 to
100 while the scattering cross section was fixed at
unity. Figure~\ref{fig:measured_spec_rad} gives the measured spectral
radius of the iteration matrix and the computed spectral radius for
the preconditioned diffusion operator using
Eq~(\ref{eq:iteration_radius}) as function of the absorption to
scattering ratio $(\Sigma_a / \Sigma_s)$. We note here that although
there is excellent agreement between the measured and analytic
results, slight differences likely arise due to the non-reentrant
current condition not being included in the spectral analysis. To
verify this (and serve as a better testing environment), reflecting
boundary conditions must be implemented.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_spec_rad.png}
  \end{center}
  \caption{\textbf{Measured and analytic preconditioned diffusion
      operator spectral radius as a function of the absorption cross
      section to scattering cross section ratio.} \textit{Values of
      $h=1$, $h=0.1$, and $h=0.01$ were used to verify the
      behavior. The red data was computed numerically by an
      Eigensolver while the black dashed data was generated by
      Eq~(\ref{eq:iteration_radius}).}}
  \label{fig:measured_spec_rad}
\end{figure}

\subsection{Random Walk Length}
With the spectral data in hand, we can go about setting up an
experiment to measure the length of the random walks generated by the
adjoint Neumann-Ulam solver. To do this, we again use a $100 \times
100$ square grid with $h=0.1$ and non-reentrant current boundary
conditions was used and the absorption cross varied from 0 to 100
while the scattering cross section was fixed at unity. A weight cutoff
of \sn{1}{-4} was used with 10,000 histories generated by a point
source source of strength 1 in the center of the domain such that the
vacuum boundary conditions will have less of an effect on the length
of the random walks. For each of the histories, the number of
transitions made was tallied to provide an effective value of $k$ for
each history. This value was then averaged over all histories to get a
measured value of $k$ for the particular
operator. Figure~\ref{fig:measured_length} gives these measurements as
well as the analytic result computed by Eq~(\ref{eq:analytic_k}) as a
function of the iteration matrix spectral radius,
$\rho(\ve{H})$. Again, we note here that although there is excellent
agreement between the measured and analytic results, slight
differences likely arise due to the non-reentrant current condition
not being included in the spectral analysis.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_length.png}
  \end{center}
  \caption{\textbf{Measured and analytic random walk length as a
      function of the iteration matrix spectral radius.} \textit{The
      weight cutoff was varied with \sn{1}{-2}, \sn{1}{-4}, and
      \sn{1}{-8}, to verify the validity of that parameter in
      Eq~(\ref{eq:analytic_k}). The red data was computed numerically
      by an adjoint Neumann-Ulam implementation while the black dashed
      data was generated by Eq~(\ref{eq:analytic_k}).}}
  \label{fig:measured_length}
\end{figure}

@@

To compute the leakage fraction numerically, over the global domain
\sn{1}{5} histories were sampled from a uniform source of strength
unity. At the start of a stage of histories, the number of histories
starting in the domain was computed and as the stage progressed, the
number of histories that exited the domain was tallied with the ratio
of the two numbers providing a numerical measure for the leakage
fraction. Figure~\ref{fig:measured_length} gives the domain leakage
measurements for the domain in the center of the global grid as well
as the analytic result computed by
Eq~(\ref{eq:analytic_domain_leakage}) as a function of the absorption
cross section. Again, we note good qualitative agreement between the
measured and analytic quantities.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{measured_leakage.png}
  \end{center}
  \caption{\textbf{Measured and analytic domain leakage as a function
      of the iteration matrix spectral radius.} \textit{To test the
      behavior with respect to domain size, $n_i=5$, $n_i=10$,
      $n_i=20$, and $n_i=100$ were used. The red data was computed
      numerically by a domain-decomposed adjoint Neumann-Ulam
      implementation while the black dashed data was generated by
      Eq~(\ref{eq:analytic_domain_leakage}).}}
  \label{fig:measured_leakage.}
\end{figure}
